An assessment of photometric redshift PDF performance in the context of LSST

### 2018-08-21 paper/tex/methods.tex
started fixing minor typos

%\red{The questions that must be answered for each code are\dots}
The questions that must be answered for each code are: what unique features are included in the specific implementation that influence the output $p(z)$.  What form of validation was performed with the training data, how were photometric uncertainties employed in the analysis, how were negative fluxes treated, what specific prior form was employed (for template based codes), or what specific machine learning architcture was used (for ML codes)?
%\scc{which parameter (in a general meaning) have been used, which kind of validation has been performed (if any)}
\subsection{Template-based Approaches}
\label{sec:templatecodes}

@@

In this paper we use \textsc{BPZ v 1.99.3}. The template set employed here is the set of $100$ discrete SEDs described in Section~\ref{sec:buzztemplates}
%, and the $129$ Brown SEDs for the Galacticus dataset.
To keep the number of free parameters to a manageable level the SEDs in the training set are sorted by the rest-frame $u$-$g$ colour and split into three ``broad'' SED classes, equivalent to the \texttt{E}, \texttt{Sp} and \texttt{Im/SB} types in \citet{Benitez:00}. We assume the same functional form for the Bayesian priors as used by \citet{Benitez:00}, and utilize the training-set galaxies with known SED-type, redshift, and apparent magnitude to determine the type fractions and the best fit for the eleven free parameters of the prior.
%For photo-$z$ point estimates we use the \texttt{Z\_B} output parameter.
For galaxies with negative flux in a measured band, the placeholder value is replaced with an estimate one $\sigma$ detection limit in that particular band, i.~e.~a value close to the estimated sky noise threshold.
The type-marginalized $p(z)$ is generated by setting the parameter \texttt{PROBS\_LITE=TRUE} in the \textsc{BPZ} parameter file.

@@

Each SED is redshifted in steps of $\Delta z = 0.01$ and convolved with the simulated LSST filter transmission curves (accounting for instrument efficiency). The opacity of the inter-galactic medium has been set to zero as no additional reddening has been included in the Buzzard simulations. The computed photo-$z$ is then the value that minimizes the merit function $\chi^2 (z,T,A)$ from \citet{Arnouts:99}:

@@

\textsc{ANNz2} is capable of producing both photo-$z$ point estimates and redshift posterior probability distributions $p(z)$, it could also conduct classifications and supports reweighting between samples. The PDFs are produced by propagating the intrinsic uncertainty on the input parameters and the uncertainty in the machine learning method to the expected photo-$z$ solution, averaged over multiple runs weighted based on the performance of each run. \textsc{ANNz2} presents its photo-$z$ uncertainty different from many codes by using the KNN method: it estimates the photo-$z$ bias between each object and a fixed number of nearest neighbours in parameter space, it then takes the $68$th percentile width of the distribution of the bias. This is based on the implication that objects with similar photometric properties would have similar uncertainties, and therefore the photometric errors of the inputs are not propagated into the code.

@@

The nearest-neighbours color-matching photometric redshift estimator (\textsc{CMNN}) is presented in \cite[][herafter G18]{Graham:17}. This method uses a training set of galaxies with known redshifts that has equivalent or better photometry as the test set in terms of quality and filter coverage. For each galaxy in the test set we identify a color-matched subset of training galaxies, choose one (e.g. the nearest-neighbour or a random selection), and use its known redshift as the photo-$z$. This color-matched subset is identified by first calculating the Mahalanobis distance $D_M$ in color-space between the test galaxy and all training-set galaxies: the difference between the test and a training set galaxy's color divided by the photometric error, summed over all colors (i.e., $u$-$g$, $g$-$r$, $r$-$i$, $i$-$z$, and $z$-$y$). Then, the threshold value for $D_M$ that define a good color match is set by the percent point function (PPF): for example, for $N_{\rm dof}=5$, PPF$=95$ per cent of all training galaxies consistent with the test galaxy will have $D_M < 11.07$ (where $N_{\rm dof}$, the number of degrees of freedom, is the number of colors). For a given test galaxy, the $p(z)$ is the normalized distribution of the true catalogue redshifts of this color-matched subset of training galaxies, and the standard deviation of the color-matched subset is used as the photo-$z$ uncertainty.

@@

In this study, all $100$ ordered Buzzard templates, as described in Section~\ref{sec:buzztemplates}, were used in \textsc{Delight}, and the Gaussian process was trained with a subset of $50\,000$ galaxies. Photometric uncertainties from the inputs are propagated into the code, while non-detections for each band are set to the mean of the respective bands. Default settings of \textsc{Delight} were use, with the exception that the PDF bins were set to be linear instead of logarithmic, with $200$ equally-spaced bins between $0.0<z<2.0$. In this study a flat prior is assumed.

@@

\subsubsection{METAPhoR}
\label{sec:metaphor}
%(Stefano Cavuoti, Massimo Brescia, Giuseppe Longo)
%\textcolor{brown}{general review done on this section}
\textsc{METAPhoR} \citep[Machine-learning Estimation Tool for Accurate Photometric Redshifts,][]{Cavuoti:17} is a pipeline designed to provide photo-$z$'s point estimates and a reliable PDF for machine learning (ML) based techniques. It includes pre- and post-processing phases, hosting a photo-$z$ prediction engine based on the Multi Layer Perceptron with Quasi Newton Algorithm (MLPQNA), already validated on photo-$z$'s in several cases \citep{de_Jong:17,Cavuoti:17b,Cavuoti:15,Brescia:14,Brescia:13,Biviano:13}. Due to its plug-in based modular nature, \textsc{METAPhoR} can be easily replaced by any other photo-$z$ prediction kernel, regardless its implementation, by taking the I/O interface compliance as unique constrain.

>>>>

%\red{The questions that must be answered for each code are\dots}
The questions that must be answered for each code are: what unique features are included in the specific implementation that influence the output $p(z)$.  What form of validation was performed with the training data, how were photometric uncertainties employed in the analysis, how were negative fluxes treated, what specific prior form was employed (for template based codes), or what specific machine learning architecture was used (for ML codes)?
%\scc{which parameter (in a general meaning) have been used, which kind of validation has been performed (if any)}
\subsection{Template-based Approaches}
\label{sec:templatecodes}

@@

In this paper we use \textsc{BPZ v 1.99.3}. The template set employed here is the set of $100$ discrete SEDs described in Section~\ref{sec:buzztemplates}
%, and the $129$ Brown SEDs for the Galacticus dataset.
To keep the number of free parameters to a manageable level the SEDs in the training set are sorted by the rest-frame $u$-$g$ colour and split into three ``broad'' SED classes, equivalent to the \texttt{E}, \texttt{Sp} and \texttt{Im/SB} types in \citet{Benitez:00}. We assume the same functional form for the Bayesian priors as used by \citet{Benitez:00}, and utilize the training-set galaxies with known SED-type, redshift, and apparent magnitude to determine the type fractions and the best fit for the eleven free parameters of the prior.
%For photo-$z$ point estimates we use the \texttt{Z\_B} output parameter.
For galaxies that are not detected in a measured band, the placeholder value is replaced with an estimate of the one $\sigma$ detection limit in that particular band, i.~e.~a value close to the estimated sky noise threshold.
The type-marginalized $p(z)$ is generated by setting the parameter \texttt{PROBS\_LITE=TRUE} in the \textsc{BPZ} parameter file.

@@

Each SED is redshifted in steps of $\Delta z = 0.01$ and convolved with the simulated LSST filter transmission curves (accounting for instrument efficiency).  The computed photo-$z$ is then the value that minimizes the merit function $\chi^2 (z,T,A)$ from \citet{Arnouts:99}:

@@

\textsc{ANNz2} is capable of producing both photo-$z$ point estimates and redshift posterior probability distributions $p(z)$.  It can also perform classifications, and supports reweighting between samples. The PDFs are produced by propagating the intrinsic uncertainty on the input parameters and the uncertainty in the machine learning method to the expected photo-$z$ solution, averaged over multiple runs weighted based on the performance of each run. \textsc{ANNz2} presents its photo-$z$ uncertainty different from many codes by using the KNN method: it estimates the photo-$z$ bias between each object and a fixed number of nearest neighbours in parameter space, it then takes the $68$th percentile width of the distribution of the bias. This is based on the implication that objects with similar photometric properties would have similar uncertainties, and therefore the photometric errors of the inputs are not propagated into the code.

@@

The nearest-neighbours color-matching photometric redshift estimator (\textsc{CMNN}) is presented in \citet[][herafter G18]{Graham:17}. This method uses a training set of galaxies with known redshifts that has equivalent or better photometry as the test set in terms of quality and filter coverage. For each galaxy in the test set we identify a color-matched subset of training galaxies, choose one (e.g. the nearest-neighbour or a random selection), and use its known redshift as the photo-$z$. This color-matched subset is identified by first calculating the Mahalanobis distance $D_M$ in color-space between the test galaxy and all training-set galaxies: the difference between the test and a training set galaxy's color divided by the photometric error, summed over all colors (i.e., $u$-$g$, $g$-$r$, $r$-$i$, $i$-$z$, and $z$-$y$). Then, the threshold value for $D_M$ that define a good color match is set by the percent point function (PPF): for example, for $N_{\rm dof}=5$, PPF$=95$ per cent of all training galaxies consistent with the test galaxy will have $D_M < 11.07$ (where $N_{\rm dof}$, the number of degrees of freedom, is the number of colors). For a given test galaxy, the $p(z)$ is the normalized distribution of the true catalogue redshifts of this color-matched subset of training galaxies, and the standard deviation of the color-matched subset is used as the photo-$z$ uncertainty.

@@

In this study, all $100$ ordered Buzzard templates, as described in Section~\ref{sec:buzztemplates}, were used in \textsc{Delight}, and the Gaussian process was trained using the provided training sample. Photometric uncertainties from the inputs are propagated into the code, while non-detections for each band are set to the mean of the respective bands. The default settings of \textsc{Delight} were used, with the exception that the PDF bins were set to be linear instead of logarithmic, with $200$ equally-spaced bins between $0.0<z<2.0$. In this study a flat prior is assumed.

@@

\subsubsection{METAPhoR}
\label{sec:metaphor}
%(Stefano Cavuoti, Massimo Brescia, Giuseppe Longo)
%\textcolor{brown}{general review done on this section}
\textsc{METAPhoR} \citep[Machine-learning Estimation Tool for Accurate Photometric Redshifts,][]{Cavuoti:17} is a pipeline designed to provide photo-$z$ point estimates and a reliable PDF for machine learning (ML) based techniques. It includes pre- and post-processing phases, hosting a photo-$z$ prediction engine based on the Multi Layer Perceptron with Quasi Newton Algorithm (MLPQNA), already validated on photo-$z$'s in several cases \citep{de_Jong:17,Cavuoti:17b,Cavuoti:15,Brescia:14,Brescia:13,Biviano:13}. Due to its plug-in based modular nature, \textsc{METAPhoR} can be easily replaced by any other photo-$z$ prediction kernel, regardless of specific implementation, while still taking advantage of the overall code interface.
